The \proglang{R} Package \pkg{forestinventory}: Design-Based Global and Small Area Estimations for Multi-Phase Forest Inventories

### 2017-09-08 rnw/Special Cases and Scenarios.Rnw
update Andreas 08092017

To demonstrate \textit{poststratification}, we first create an artificial categorical variable \textit{developement stage} (\code{stage}) by clustering the mean canopy heights of the \code{grisons} data set into 3 classes:

@@

\textit{Two-phase sampling for regression within strata} implies the combination of continuous and categorical variables within the model (i.e. we have an ANCOVA model). If an interaction term is not present between categorical and continuous variables, the regression lines within the strata will have the same slope but different intercepts. If an interaction term is present, both the intercept and the slope are allowed to vary within the strata. Note that one can actually use the entire range of ols regression techniques in the multi-phase estimators.

@@

As mentioned in section \ref{sec:packstruc}, \textit{cluster sampling} is a special case of sample designs where the sample location, i.e. sample unit, does no longer consist of one, but multiple spatially agglomerated sample points. Whereas the maximal possible number of sample points per cluster $M$ is defined by the user, the actual number of points $M(x)$ falling into the forest area $F$ at sample location $x$ is random. Without going into too much mathematical detail, the estimators under simple sampling are extended in a way that all parameters (local density, mean vector of explanatory variables, mean model residuals) have to be calculated as the weighted cluster means with $M(x)$ being the cluster weights (for more detail see Appendix). The package functions allow to define cluster-structures by specifying the argument \code{cluster}.

>>>>

To demonstrate \textit{poststratification}, we first create an artificial categorical variable \textit{developement stage} (\code{stage}) by clustering the mean canopy heights of the \code{grisons} data set into 3 height classes:

@@

\textit{Two-phase sampling for regression within strata} implies the combination of continuous and categorical variables within the model (i.e. we have an ANCOVA model). If an interaction term is not present between categorical and continuous variables, the regression lines within the strata will have the same slope but different intercepts. If an interaction term is present, both the intercept and the slope are allowed to vary within the strata. Note that one can actually use the entire range of ols regression techniques in the multi-phase estimators, including higher order terms and transformations of the explanatory variables.

@@

As mentioned in section \ref{sec:packstruc}, \textit{cluster sampling} is a special case of sample designs where the sample unit does no longer consist of one, but multiple spatially agglomerated sample points. One still randomly places the sample location $x$ in the inventory area likewise in the simple sampling design, but then creates $M-1$ additional sample locations $x_2,...,x_M$ close to the cluster origin $x$ by adding a fixed set of spatial vectors $e_2,...,e_M$ to $x$. The idea of cluster sampling is to increase the amount of information without increasing the travelling costs of the terrestrial campaign. Not however that the information gathered at all sub-locations of a cluster is then averaged on the cluster level, and that this average value is again referred to exactly one point, i.e. the cluster origin $x$, in the frame of the infinite population approach. Without going into too much mathematical detail, the estimators under simple sampling are thus extended in a way that all parameters (local density, mean vector of explanatory variables, mean model residuals) have to be calculated as the weighted cluster means with $M(x)$ being the cluster weights (for more detail see Appendix). Whereas the geometric form and the number of sample locations per cluster $M$ is fixed (i.e defined by the user), the actual number of points $M(x)$ falling into the forest area $F$ at sample location $x$ is random because the cluster origin $x$ is random. The package functions of \pkg{forstinventory} also allow to consider cluster-structures by specifying a unique cluster identification (cluster ID) for each cluster in an additional column whose column name is passed to the argument \code{cluster} in the \code{twophase()} and \code{threephase()} function calls.
